---
title: "exercise2"
author: "Valerie Arnaldi"
date: "29 4 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the necessary libraries ################################################
```{r}
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times
```
## Task 1 #####################################################################

## Import the downloaded csv ##################################################
```{r}
wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",delim=",",locale=locale
                             (encoding = "UTF-8"))

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056,
                           remove = FALSE)
```
# dataset is already converted to EPSG 2056
# coordinates are stored in the columns (E/N)
# setting remove = FALSE preserves the original (E/N) columns


## Task 2 #####################################################################

## Getting an overview ########################################################

# Calculate the time difference between subsequent rows

```{r}
wildschwein <- mutate(wildschwein_BE,timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "sec")))

wildschwein
```
# How many individuals were tracked?
```{r}
individuals <- wildschwein %>% group_by(TierName) %>%
  summarize(count=n())
nrow(individuals)
```
# 3 individuals were tracked

# For how long were the individual tracked? Are there gaps?
```{r}
ggplot(wildschwein_BE, aes(DatetimeUTC,TierName))+
  geom_line()
```
# Sabi from mid 2014 until mid 2015
# Ruth from end 2014 until mid 2015
# Rosa from end 2014 until july 2015
# without gaps

# Were all individuals tracked concurrently or sequentially?
```{r}
ggplot(wildschwein, aes(timelag))+
  geom_bar()+
  scale_y_log10()+
  scale_x_continuous(limits = c(0,15000))
```
# muessen wir da etwas mit den NA Werten machen?

# What is the temporal sampling interval between the locations?
```{r}
ggplot(wildschwein, aes(DatetimeUTC,timelag, color=TierID))+
  geom_line()+
  geom_point()+
  scale_x_datetime(date_breaks="1 month", date_labels="%b")+
  scale_y_continuous(limits = c(0,20000))
```
# nur fuer Zeit Sept bis Jan auswaehlen

## Task 3 #####################################################################

## Deriving movement parameters I: Speed ######################################

# calculate the animal’s steplength based on the Euclidean distance between two
# subsequent locations
```{r}
wildschwein <- wildschwein %>%
  mutate(steplength = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2)
  )
```
# in welcher Einheit wird die Schrittlaenge ausgegeben?

# Calculate the animals speed between consecutive locations based on steplength and timelag
```{r}
wildschwein <- wildschwein %>%
  mutate(speed = steplength / timelag)
```
# Einheit kommt darauf an, was die Ausgabe der Schrittlaenge ist. Das andere sind Sekunden (vgl. timelag)

## Task 4 #####################################################################

## Cross-scale movement analysis #############################################

# analyse animal movement across different scales
# reducing the granularity of the data by subsetting the data to every nth element
# 200 locations of a single wild boar with a constant sampling interval of 60 seconds

# import Data
```{r}
caro <- read_delim("caro60.csv",delim=",")

caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)
```
# the locations are stored in EPSG 2056

# reduce granularity
```{r}
caro_3 <- caro %>%
  slice(seq(from = 1, to = 200, by = 3))  #'slice' nimmt Anzahl Zeilen heraus, 'seq' gibt ihm die Wiederholung aller drei Zeilen
  
caro_6 <- caro %>%
  slice(seq(from = 1, to = 200, by = 6))

caro_9 <- caro %>%
  slice(seq(from = 1, to = 200, by = 9))
```
# control
```{r}
nrow(caro) #200
nrow(caro_3) #67
nrow(caro_6) #34
nrow(caro_9) #23
```
# calculate timelag, steplength, speed for each data set

# timelag
```{r}
caro <- mutate(caro, timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "sec")))

caro_3 <- mutate(caro_3, timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "sec")))

caro_6 <- mutate(caro_6, timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "sec")))

caro_9 <- mutate(caro_9, timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "sec")))
```
# steplength
```{r}
caro <- caro %>%
  mutate(steplength = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2)
  )

caro_3 <- caro_3 %>%
  mutate(steplength = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2)
  )

caro_6 <- caro_6 %>%
  mutate(steplength = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2)
  )

caro_9 <- caro_9 %>%
  mutate(steplength = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2)
  )
```
# speed
```{r}
caro <- caro %>%
  mutate(speed = steplength / timelag)

caro_3 <- caro_3 %>%
  mutate(speed = steplength / timelag)

caro_6 <- caro_6 %>%
  mutate(speed = steplength / timelag)

caro_9 <- caro_9 %>%
  mutate(speed = steplength / timelag)
```
# visualize the trajectories in a map
```{r}
ggplot()+
  geom_path(caro, mapping=aes(E,N, color="1 minute"))+
  geom_point(caro, mapping=aes(E,N, color="1 minute"))+
  geom_path(caro_3, mapping=aes(E,N, color="3 minute"))+
  geom_point(caro_3, mapping=aes(E,N, color="3 minute"))+
  labs(title="Comparing original with 3-minute-resampling",
       color="Trajectory")

ggplot()+
  geom_path(caro, mapping=aes(E,N, color="1 minute"))+
  geom_point(caro, mapping=aes(E,N, color="1 minute"))+
  geom_path(caro_6, mapping=aes(E,N, color="6 minute"))+
  geom_point(caro_6, mapping=aes(E,N, color="6 minute"))+
  labs(title="Comparing original with 6-minute-resampling",
       color="Trajectory")

ggplot()+
  geom_path(caro, mapping=aes(E,N, color="1 minute"))+
  geom_point(caro, mapping=aes(E,N, color="1 minute"))+
  geom_path(caro_9, mapping=aes(E,N, color="9 minute"))+
  geom_point(caro_9, mapping=aes(E,N, color="9 minute"))+
  labs(title="Comparing original with 9-minute-resampling",
       color="Trajectory")
```
# Comparing derived speed at different sampling intervals
```{r}
ggplot()+
  geom_line(caro, mapping=aes(DatetimeUTC,speed, color="1 minute"))+
  geom_line(caro_3, mapping=aes(DatetimeUTC,speed, color="3 minutes"))+
  geom_line(caro_6, mapping=aes(DatetimeUTC,speed, color="6 minutes"))+
  geom_line(caro_9, mapping=aes(DatetimeUTC,speed, color="9 minutes"))+
  labs(title="Comparing derived speed at different sampling intervals",
       color="Interval",
       x="Time",
       y="speed (m/s)")
```

## Task 5 #####################################################################

## Deriving movement parameters II: Rolling window functions ##################

```{r}
install.packages("roll_*")
library(zoo)
```
# example
```{r}
example <- rnorm(10)

rollmean(example,k = 3,fill = NA,align = "left")

rollmean(example,k = 4,fill = NA,align = "left")
```
# run rollmean on the speed variable of the subset (caro)
```{r}
rollmean(caro$speed,k = 3,fill = NA,align = "left")
rollmean(caro_3$speed,k = 3,fill = NA,align = "left")
rollmean(caro_6$speed,k = 3,fill = NA,align = "left")
rollmean(caro_9$speed,k = 3,fill = NA,align = "left")
```
# visualize with different window sizes
```{r}
caro <- caro %>%
  mutate(rollmean3 = rollmean(caro$speed,k = 3,fill = NA,align = "left"))

caro <- caro %>%
  mutate(rollmean5 = rollmean(caro$speed,k = 5,fill = NA,align = "left"))

caro <- caro %>%
  mutate(rollmean8 = rollmean(caro$speed,k = 8,fill = NA,align = "left"))

ggplot()+
  geom_line(caro, mapping=aes(x=DatetimeUTC, y=rollmean3, color="k=3"))+
  geom_line(caro, mapping=aes(x=DatetimeUTC, y=rollmean5, color="k=5"))+
  geom_line(caro, mapping=aes(x=DatetimeUTC, y=rollmean8, color="k=8"))+
  labs(title="Comparing derived speed by rolling window functons",
       color="Window Sizes",
       x="Time",
       y="speed (m/s)")
```



